A Two-Infall Star Formation History with Outflows and Radial Migration

### 2025-02-12 src/tex/ms.tex
Updates from Overleaf

% The parameters of Equation \ref{eq:infall-rate} have important effects on the chemical evolution of the model. 
% Previous studies of the two-infall model have adopted values for the infall parameters of $\tau_1\sim0.1-1$ Gyr, $\tau_2\sim3-10$ Gyr, and $t_{\rm max}\sim3-5$ Gyr. 
Figure \ref{fig:twoinfall-parameters} illustrates the effect of varying the infall parameters on gas abundance tracks and stellar abundance distributions in a one-zone model. We note that all our models are normalized to produce the same thick-to-thin-disk mass ratio. The first infall timescale $\tau_1$, shown in panel (a), primarily affects the stellar distribution along the high-$\alpha$ sequence. Though $\tau_1$ has an apparently large effect on the size of the low-$\alpha$ loop, the effect on the stellar abundance distribution of the low-$\alpha$ sequence is quite small due to the low number of stars formed between $t\sim3-6$ Gyr. We adopt $\tau_1=1$ Gyr for our fiducial value, in line with \citet{spitoni_galactic_2020} but longer than, e.g., \citet{nissen_high-precision_2020} or \citet{spitoni_apogee_2021}, in order to set the peak of the high-$\alpha$ sequence at ${\rm [O/Fe]}\approx+0.3$.

@@

@@

\subsection{Stellar Abundance Distributions}
\label{sec:abundance-distributions}

@@

\subsection{Age--Abundance Relations}
\label{sec:age-abundance}

@@

@@

@@

A bursty infall history is not the only way to produce a gap in the star formation history. \citep{beane_rising_2024} present

\begin{figure}
    \centering
    \includegraphics{figures/onezone_sfe_hiatus.pdf}
    \caption{Abundance tracks and distributions from one-zone models which experience an efficiency-driven starburst. The blue dashed curve represents the fiducial model that has an exponentially declining infall rate and constant star formation efficiency timescale $\tau_\star=2$ Gyr. The red solid curve plots the output of a model which experiences an enhancement of $\tau_\star$ by a factor of 10, for a duration of 200 Myr, starting at $t=1.4$ Gyr. The greyscale histogram presents the number density of APOGEE stars in [O/Fe]--[Fe/H] space, and the gray histograms in the marginal panels show the APOGEE stellar abundance distributions.}
    \label{fig:onezone-sfe-hiatus}
\end{figure}

>>>>

% The parameters of Equation \ref{eq:infall-rate} have important effects on the chemical evolution of the model. 
% Previous studies of the two-infall model have adopted values for the infall parameters of $\tau_1\sim0.1-1$ Gyr, $\tau_2\sim3-10$ Gyr, and $t_{\rm max}\sim3-5$ Gyr. 
Figure \ref{fig:twoinfall-parameters} illustrates the effect of varying the infall parameters on gas abundance tracks and stellar abundance distributions in a one-zone model. The first infall timescale $\tau_1$, shown in panel (a), primarily affects the stellar distribution along the high-$\alpha$ sequence. Though $\tau_1$ has an apparently large effect on the size of the low-$\alpha$ loop, the effect on the stellar abundance distribution of the low-$\alpha$ sequence is quite small due to the low number of stars formed between $t\sim3-6$ Gyr. We adopt $\tau_1=1$ Gyr for our fiducial value, in line with \citet{spitoni_galactic_2020} but longer than, e.g., \citet{nissen_high-precision_2020} or \citet{spitoni_apogee_2021}, in order to set the peak of the high-$\alpha$ sequence at ${\rm [O/Fe]}\approx+0.3$.

@@

We note that all our models are normalized to produce the same thick-to-thin-disk mass ratio regardless of the infall parameters. The high-$\alpha$ sequence appears much less prominent in our [O/Fe] distributions in Figure \ref{fig:twoinfall-parameters} than in the data because the model outputs include only stars which were formed in-situ at the Solar annulus. In our multi-zone models, most of the high-$\alpha$ stars present in the Solar neighborhood have migrated from the inner Galaxy.

@@

\subsection{Present-Day Abundances}
\label{sec:abundance-distributions}

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_density.pdf}
    \caption{The density of stars in the [O/Fe]--[Fe/H] plane from multi-zone models with (a) $y/Z_\odot=1$ and (b) $y/Z_\odot=2$, and (c) from the APOGEE DR17 catalog. Stars are restricted to the region defined by $7\leq R_{\rm gal}< 9$ kpc and $0.5\leq|z|<1$ kpc to highlight the $\alpha$-bimodality.}
    \label{fig:ofe-feh-density}
\end{figure}

Figure \ref{fig:ofe-feh-density} illustrates the origin of the intermediate-$\alpha$ peak predicted by the two-infall model at mid to high Galactic latitudes. Between $0.5\leq|z|<1$ kpc, both the models with $y/Z_\odot=1$ and $y/Z_\odot=2$ predict an over-density of stars near the abundance turn-over (${\rm [Fe/H]}\approx-0.3$, ${\rm [O/Fe]}\approx0.1-0.2$), which is not seen in the APOGEE sample. Additionally, it is clear that the shape of the low-$\alpha$ sequence in the model results (a concave-down ``comma'') is different from the data (a concave-up ``swoosh'').

\subsection{Abundance Evolution}
\label{sec:age-abundance}

@@

\subsection{Abundances Across the Disk}
\label{sec:disk-abundances}

@@

\begin{figure}
    \centering
    \includegraphics{figures/onezone_sfe_hiatus.pdf}
    \caption{Abundance tracks and distributions from one-zone models which experience an efficiency-driven starburst. The blue dashed curve represents the fiducial model that has an exponentially declining infall rate and constant star formation efficiency timescale $\tau_\star=2$ Gyr. The red solid curve plots the output of a model which experiences an enhancement of $\tau_\star$ by a factor of 10, for a duration of 200 Myr, starting at $t=1.4$ Gyr. Both models assume the {\tt yZ2} yield set, with $y_{\rm Fe}^{\rm Ia}$ reduced by 20\% to better match the model endpoint with the data, and $\eta=1.4$. The greyscale histogram presents the number density of APOGEE stars in the Solar annulus ($7\leq R_{\rm gal}\leq 9$ kpc, $0\leq|z|\leq2$ kpc) in [O/Fe]--[Fe/H] space, and the gray histograms in the marginal panels show the APOGEE stellar abundance distributions.}
    \label{fig:onezone-sfe-hiatus}
\end{figure}

@@

A bursty infall history is not the only way to produce a gap in the star formation history. \citet{beane_rising_2024} present a simulated galaxy from the Illustris TNG50 suite that exhibits MW-like bimodality. They argue that the $\alpha$-bimodality is brought on by a brief ($\sim300$ Myr) quiescent period caused by bar formation. The virial mass of their galaxy grows steadily throughout this period, unlike in our two-infall model where the mass grows by a factor of \todo{X} during the 1 Gyr following the second infall.

While our semi-analytic model does not include a Galactic bar, we can explore the effects of a star formation hiatus by artificially boosting the SFE timescale $\tau_\star$ for a period of time. Figure \ref{fig:onezone-sfe-hiatus} illustrates the effect of this SFE-driven hiatus in a one-zone model with an exponentially declining infall rate. During the quiescent period, the [O/Fe] ratio slowly declines due to the delayed contribution of Fe from SNe Ia. Meanwhile, the gas mass continues to increase even as star formation is suppressed. When $\tau_\star$ is lowered at the end of the quiescent period, the high gas mass sparks a moderate star formation burst which causes stellar abundances to ``pile up'' at similar [O/Fe] values. The trough between the high- and low-$\alpha$ sequences results from the star formation returning to pre-quiescence behavior.

Our simple hiatus model offers a few parameters which control the chemical evolution. The onset time of the SFE hiatus controls the position of the high-$\alpha$ sequence: a later onset places the peak at lower [O/Fe]. The duration of the star formation hiatus \todo{(and the $\tau_\star$ enhancement factor?)} controls the strength of the high-$\alpha$ peak.

The parameters of the SFE hiatus in Figure \ref{fig:onezone-sfe-hiatus} were chosen to match the APOGEE stellar [O/Fe] distribution as closely as possible. However, there are some differences in detail, such as the dearth of stars at ${\rm [O/Fe]}\approx+0.35$ due to the star formation hiatus. We intend this model to illustrate another path to reproducing the $\alpha$-bimodality. Most of the high-$\alpha$ stars present in the Solar annulus have likely migrated from the inner Galaxy, where perhaps this SFE-driven hiatus was concentrated.
